---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  fig.height = 3
)
```

# meteostatR

<!-- badges: start -->
`r badger::badge_devel("jmartindelasierra/meteostatR", "blue")`
<!-- badges: end -->

This package provides a set of functions to retrieve historical weather data from [Meteostat](https://meteostat.net), a free online weather and climate service, by connecting to its JSON API.

I encourage you to visit [Meteostat](https://meteostat.net) and, especially, the developer's [site](https://dev.meteostat.net/). Please, be sure you understand the [terms and license](https://dev.meteostat.net/docs/terms.html) before continue.

## Installation

You can install this package as follows:

```{r eval=FALSE}
devtools::install_github("jmartindelasierra/meteostatR")
```

## Preparation

Before using the package, this requires an entry in your `.Renviron` file to store your Meteostat API key.

You can obtain your key from the Meteostat developer's [site](https://dev.meteostat.net/). Once received, edit your `.Renviron` file with `usethis::edit_r_environ()` by adding the following line:

`METEOSTAT_KEY = your_key`

Remember not to use quotes for your key string.

Next you must restart R to take effect and finally check your key was recorded:

```{r eval=FALSE}
Sys.getenv("METEOSTAT_KEY")
```

## Functions

This is the set of functions provided in the package:

- find_stations()

    Returns weather stations by the indicated string.

- nearby_stations()

    Returns weather stations nearby the indicated geolocation.

- metadata_station()

    Returns meta data for the indicated weather station.

- hourly_data()

    Returns hourly weather data from the indicated weather station.

- point_hourly_data()

    Returns hourly weather data from the indicated geolocation by means of interpolation.

- daily_data()

    Returns daily weather data from the indicated station.

- point_daily_data()

    Returns daily weather data from the indicated geolocation by means of interpolation.

- climate_normals()

    Returns long-term weather averages from the indicated station.

- point_climate_normals()

    Returns long-term weather averages from the indicated geolocation by means of interpolation.

Get more about the return values visiting the Meteostat [API](https://dev.meteostat.net/api/) documentation.

## Example

```{r include=FALSE}
# Save current locale
system_locale <- Sys.getlocale(category = "LC_TIME")

# Set time locale to english
Sys.setlocale(category = "LC_TIME", locale = "en_US.UTF-8")
```

Let's show how to use some functions with an example.

We start by attaching the required packages:

```{r message=FALSE, warning=FALSE}
# This package
library(meteostatR)

# Other packages for data manipulation and graphics
library(dplyr)
library(ggplot2)
```

We look for stations located in Barcelona:

```{r}
# Find stations including "barcelona"
stations <- find_stations("barcelona")
```

```{r}
# Look at the returned data
stations$data
```

By filtering the stations in Spain, we see that only one is active right now:

```{r}
# Filter by country of interest
stations$data %>%
  filter(country == "ES") %>%
  select(id, name, latitude, longitude, active)
```

We keep such station and retrieve some hourly data:

```{r}
# Save station of interest
my_station <- "08181"
```

```{r}
# Retrieve hourly data
hourly <- hourly_data(station = my_station, 
                      start = "2021-02-01", 
                      end = "2021-02-10")
```

Let's see what features we obtained:

```{r}
# Variables in data
hourly$data %>% 
  glimpse()
```

We may want to plot some data:

```{r}
# Plot pressure
hourly$data %>% 
  ggplot(aes(x = time, y = pres)) + 
  geom_line(color = "steelblue") + 
  theme_bw() +
  labs(x = NULL)
```

Now we're going to retrieve daily data for 2020 and plot the average temperature:

```{r}
# Retrieve daily data
daily <- daily_data(station = my_station, 
                    start = "2020-01-01", 
                    end = "2020-12-31")
```

```{r}
# Variables in data
daily$data %>%
    glimpse()
```

```{r}
# Plot the average temperature with minimum and maximum ribbon
daily$data %>%
  ggplot(aes(x = date)) + 
  geom_ribbon(aes(ymin = tmin, ymax = tmax), fill = "steelblue", alpha = 0.2) + 
  geom_line(aes(y = tavg), color = "steelblue") + 
  theme_bw() + 
  labs(x = NULL, y = "temp")
```

```{r include=FALSE}
Sys.setlocale(category = "LC_TIME", locale = system_locale)
```

## Quota

Free access to the Meteostat API is limited to 2,000 requests per day and no more than two requests per second.

## Attributions

meteostatR: [Jesús Martín de la Sierra](https://github.com/jmartindelasierra).

Source: [Meteostat](https://meteostat.net). Raw data provided by [NOAA](https://www.noaa.gov/), [DWD](https://www.dwd.de/) and [others](https://dev.meteostat.net/docs/sources.html).
